---
title: "convergence"
bibliography: references.bib
---

## What did I set out to do? Introduction

The issue of technological convergence has played a substantial role in debates regarding large-scale population demographic processes and human behavioral evolution [@clarkson2018; @mellars2006]

Hypothesis: Raw material selection should drive lithic technological convergence in environmentally diverse areas. Under this scenario, we expect significant shape and size differences to exist between flaked products.

## How did I do it? (Materials and Methods)

This study attempts to evaluate lithic technological convergence with statistical measures. We specifically

```{r}
library(tidyverse)
library(ggsci)
library(ggrepel)
```

```{r}
bpa_flakes = read.csv("C:/Users/caleb/Desktop/bpa_data/bpa_flake_data.csv")
elongation <- bpa_flakes$TechLength/bpa_flakes$MaxTechWidth
flatness <- bpa_flakes$MaxTechWidth/bpa_flakes$MidThickness
surface_area <- bpa_flakes$MaxTechWidth*bpa_flakes$TechLength
bpa_flakes_2 <- bpa_flakes %>%
  mutate(elongation, flatness, surface_area)
form <- bpa_flakes_2$elongation/bpa_flakes_2$flatness
bpa_flakes_3 <- bpa_flakes_2 %>%
  mutate(form)
bpa_flakes_3$EvidenceBipolarPercussion[bpa_flakes_3$EvidenceBipolarPercussion == 'Yes'] <- 'Bipolar'
bpa_flakes_3$EvidenceBipolarPercussion[bpa_flakes_3$EvidenceBipolarPercussion == 'No'] <- 'Freehand'
bpa_flakes_updated <- subset(bpa_flakes_3, LithicArtifactClass == 'CompFlake')
bpa_bladelets <- subset(bpa_flakes_updated, Bladelet == 'Yes')
bpa_retouched_utilized <- subset(bpa_flakes_3,Completeness == 'Complete' & LithicArtifactClass == 'RetouchedPiece' | LithicArtifactClass =='Utilized')
bpa_retouched_utilized
```

core data

```{r}
bpa_cores = read.csv("C:/Users/caleb/Desktop/bpa_data/bpa_core_data.csv")
surface_area_cores <- (2*bpa_cores$MaxLength*bpa_cores$MaxWidth) + (2*bpa_cores$MaxWidth*bpa_cores$MaxThickness) + (2*bpa_cores$MaxLength*bpa_cores$MaxThickness)
SDI= bpa_cores$OverallScarCount/surface_area_cores
bpa_cores_2 <- bpa_cores %>%
  mutate(surface_area_cores, SDI)
bpa_cores_2$EvidenceBipolar[bpa_cores_2$EvidenceBipolar == 'Yes'] <- 'Bipolar'
bpa_cores_2$EvidenceBipolar[bpa_cores_2$EvidenceBipolar == 'No'] <- 'Freehand'
bpa_cores_updated <- subset(bpa_cores_2, CoreTypology == 'Core>2flakescars')

```

Split flakes by SubMember

```{r}
CL1_flakes = subset(bpa_flakes_updated, SubMember == 'CL1')
CL2_flakes = subset(bpa_flakes_updated, SubMember == 'CL2')
CL3_flakes = subset(bpa_flakes_updated, SubMember == 'CL3')
CL4_flakes = subset(bpa_flakes_updated, SubMember == 'CL4')
```

## What I found (Results)

### Assemblage Composition

```{r}
#flakes raw material composition
bpa_flakes_updated %>%
  group_by(SubMember) %>%
  count(RawMaterial) %>%
  ggplot(aes(x = as.factor(SubMember), y = n, fill = factor(RawMaterial))) +
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Raw Material")
ggsave("fig-raw-material-composition.png")

#cores raw material composition
bpa_cores_updated %>%
  group_by(SubMember) %>%
  count(RawMaterial) %>%
  ggplot(aes(x = as.factor(SubMember), y = n, fill = factor(RawMaterial))) +
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Raw Material")
ggsave("fig-raw-material-composition_cores.png")
```

### **Convergence in Flake Reduction Strategy**

```{r}
#| label: fig-reduction-method
#| fig-cap: "CL 1-4 Bipolar and freehand reduction frequencies for all raw material types."
#| warning: false

bpa_flakes_updated %>%
  group_by(RawMaterial, SubMember) %>%
  count(EvidenceBipolarPercussion) %>%
  ggplot(aes(x = as.factor(RawMaterial), y = n, fill = EvidenceBipolarPercussion))+
  geom_col(position = "fill", colour="black")+
  coord_flip()+
  labs(x="Raw Material",y="Proportion", fill="Reduction Strategy")+ 
  facet_wrap(~SubMember)
ggsave("fig-reduction-method.png")

bipolar_freehand_flakes <- chisq.test(table(bpa_flakes_updated$RawMaterial, bpa_flakes_updated$EvidenceBipolarPercussion))
bipolar_freehand_flakes
```

Technical Category/Reduction Stage

```{r}
#| label: fig-technical-category
#| fig-cap: "CL 1-4 Technical Category frequencies for all raw material types."
#| warning: false
bpa_flakes_updated %>%
  group_by(RawMaterial, SubMember) %>%
  count(TechnicalCategory) %>%
  ggplot(aes(x = as.factor(RawMaterial), y = n, fill = factor(TechnicalCategory, levels=c("A1", "A2", "A3", "B1", "B2", "B4", "B6", "B8", "B9", "B11", "C1", "C2", "D1", "D2")))) +
  coord_flip()+
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Technical Category")+ 
  facet_wrap(~SubMember)
ggsave("fig-technical-category.png")

#| label: fig-reduction_stage
#| fig-cap: "CL 1-4 Reduction Stage frequencies for all raw material types."
#| warning: false
bpa_flakes_updated %>%
  group_by(RawMaterial, SubMember) %>%
  count(Stage) %>%
  ggplot(aes(x = as.factor(RawMaterial), y = n, fill = factor(Stage, levels=c("Initial", "Main", "Core Maintenance", "Other")))) +
  coord_flip()+
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Reduction Stage")+ 
  facet_wrap(~SubMember)
ggsave("fig-reduction-stage.png")
```

### Convergence in Flake Shape

Elongation (Flakes)

```{r}
#testing elongation data for normality
qqnorm(bpa_flakes_updated$elongation)
qqline(bpa_flakes_updated$elongation)

ggplot(bpa_flakes_updated, mapping = aes(x=elongation))+
  geom_histogram()

#plot 1 
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(elongation))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#elongation anova
elongation_anova <- aov(log(elongation) ~ as.factor(RawMaterial), data=bpa_flakes_updated)
summary(elongation_anova)

#plot 2
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(elongation))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember))+
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+
  labs(x="Raw Material", y="log elongation")
ggsave("fig-flake-elongation.png")

#elongation anova
elongation_anova_2 <- aov(log(elongation) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_flakes_updated)
summary(elongation_anova_2)

#post-hoc tests
flake_elongation_tukey = TukeyHSD(elongation_anova_2)
plot(flake_elongation_tukey)
```

Elongation (bladelets)

```{r}
#testing elongation data for normality
qqnorm(bpa_bladelets$elongation)
qqline(bpa_bladelets$elongation)
ggplot(bpa_bladelets, mapping = aes(x=elongation))+
  geom_histogram()

#plot 1 
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(elongation))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#elongation anova
bladelet_elongation_anova <- aov(log(elongation) ~ as.factor(RawMaterial), data=bpa_bladelets)
summary(bladelet_elongation_anova)

#plot 2
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(elongation))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember))+
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+
  labs(x="Raw Material", y="log elongation")
ggsave("fig-bladelet-elongation.png")

#elongation anova
bladelet_elongation_anova_2 <- aov(log(elongation) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_bladelets)
summary(bladelet_elongation_anova_2)

TukeyHSD(bladelet_elongation_anova_2)
```

Flatness (Flakes)

```{r}
#testing flatness data for normality
qqnorm(bpa_flakes_updated$flatness)
qqline(bpa_flakes_updated$flatness)
ggplot(bpa_flakes_updated, mapping = aes(x=flatness))+
  geom_histogram()

#plot 1 
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = flatness)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = flatness)) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+
  labs(x="Raw Material", y="log flatness")
ggsave("fig-flake-flatness.png")

#flake_flatness_anova
flake_flatness_anova_2 <- aov(log(flatness) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_flakes_updated)
summary(flake_flatness_anova_2)

TukeyHSD(flake_flatness_anova_2)
```

Flatness (bladelets)

```{r}
#testing elongation data for normality
qqnorm(bpa_bladelets$flatness)
qqline(bpa_bladelets$flatness)
ggplot(bpa_bladelets, mapping = aes(x=flatness))+
  geom_histogram()

#plot 1 
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(flatness))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#elongation anova
bladelet_flatness_anova <- aov(log(flatness) ~ as.factor(RawMaterial), data=bpa_bladelets)
summary(bladelet_flatness_anova)

#plot 2
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(flatness))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember))+
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1)) +
  labs(x="Raw Material", y="log flatness")
ggsave("fig-bladelet-flatness.png")

#flatness anova
bladelet_flatness_anova_2 <- aov(log(flatness) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_bladelets)
summary(bladelet_flatness_anova_2)

TukeyHSD(bladelet_flatness_anova_2)
```

Form Flakes

```{r}
#testing form data for normality
qqnorm(bpa_flakes_updated$form)
qqline(bpa_flakes_updated$form)
ggplot(bpa_flakes_updated, mapping = aes(x=form))+
  geom_histogram()

#plot 1 
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(form))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(form))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log form")
ggsave("fig-flake-form.png")

#flake_form_anova
flake_form_anova_2 <- aov(log(form) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_flakes_updated)
summary(flake_form_anova_2)

TukeyHSD(flake_form_anova_2)
```

Form (bladelets)

```{r}
#testing elongation data for normality
qqnorm(bpa_bladelets$form)
qqline(bpa_bladelets$form)
ggplot(bpa_bladelets, mapping = aes(x=form))+
  geom_histogram()

#plot 1 
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(form))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#form anova
bladelet_form_anova <- aov(log(form) ~ as.factor(RawMaterial), data=bpa_bladelets)
summary(form_anova)

#plot 2
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(form))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember))+
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+
  labs(x="Raw Material", y="log form")
ggsave("fig-bladelet_form.png")

#form anova
bladelet_form_anova_2 <- aov(log(form) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_bladelets)
summary(bladelet_form_anova_2)

TukeyHSD(bladelet_form_anova_2)
```

### **Convergence in Flake Size**

Surface Area

```{r}
#testing form data for normality
qqnorm(bpa_flakes_updated$surface_area)
qqline(bpa_flakes_updated$surface_area)
ggplot(bpa_flakes_updated, mapping = aes(x=surface_area))+
  geom_histogram()

#plot 1 
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(surface_area))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(surface_area))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))
ggsave("bpa_flake_sa.png")

##anova
flake_sa_anova <- aov(log(surface_area) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_flakes_updated)
summary(flake_sa_anova)

TukeyHSD(flake_sa_anova)
```

Mass

```{r}
#normality check
qqnorm(bpa_flakes_updated$Mass)
qqline(bpa_flakes_updated$Mass)
ggplot(bpa_flakes_updated, mapping = aes(x=Mass))+
  geom_histogram()

#plot 1 
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(Mass))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_flakes_updated, aes(x= as.factor(RawMaterial), y = log(Mass))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))
ggsave("fig_flake_mass.png")

##anova
flake_mass_anova <- aov(log(Mass) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_flakes_updated)

TukeyHSD(flake_mass_anova)
```

bladelet surface area

```{r}
#testing form data for normality
qqnorm(bpa_bladelets$surface_area)
qqline(bpa_bladelets$surface_area)
ggplot(bpa_bladelets, mapping = aes(x=surface_area))+
  geom_histogram()

#plot 1 
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(surface_area))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_bladelets, aes(x= as.factor(RawMaterial), y = log(surface_area))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))
ggsave("bpa_bladelet_sa.png")

##anova
bladelet_sa_anova <- aov(log(surface_area) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_bladelets)
summary(bladelet_sa_anova)

TukeyHSD(bladelet_sa_anova)
```

### Retouched/Utilized pieces

Elongation

```{r retouched elongation}
#normality check
qqnorm(bpa_retouched_utilized$elongation)
qqline(bpa_retouched_utilized$elongation)
ggplot(bpa_retouched_utilized, mapping = aes(x=elongation))+
  geom_histogram()

#plot 1 
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(elongation))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(elongation))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1)) + 
  labs(x="Raw Material", y="log elongation")
ggsave("fig_retouch_elongation.png")

##anova
retouched_elongation_anova <- aov(log(elongation) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_retouched_utilized)
summary(retouched_elongation_anova)

TukeyHSD(retouched_elongation_anova)
```

Flatness

```{r retouched flatness}
#normality check
qqnorm(bpa_retouched_utilized$flatness)
qqline(bpa_retouched_utilized$flatness)
ggplot(bpa_retouched_utilized, mapping = aes(x=flatness))+
  geom_histogram()

#plot 1 
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(flatness))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(flatness))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1)) +
  labs(x="Raw Material", y="log flatness")
ggsave("fig_retouch_flatness.png")

##anova
retouched_flatness_anova <- aov(log(flatness) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_retouched_utilized)
summary(retouched_flatness_anova)

TukeyHSD(retouched_flatness_anova)
```

Form

```{r surface area}
#normality check
qqnorm(bpa_retouched_utilized$form)
qqline(bpa_retouched_utilized$form)
ggplot(bpa_retouched_utilized, mapping = aes(x=form))+
  geom_histogram()

#plot 1 
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(form))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(form))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+
  labs(x="Raw Material", y="log form")
ggsave("fig_retouch_form.png")

##anova
retouched_form_anova <- aov(log(form) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_retouched_utilized)
summary(retouched_form_anova)

TukeyHSD(retouched_form_anova)
```

Surface area

```{r surface area}
#normality check
qqnorm(bpa_retouched_utilized$surface_area)
qqline(bpa_retouched_utilized$surface_area)
ggplot(bpa_retouched_utilized, mapping = aes(x=surface_area))+
  geom_histogram()

#plot 1 
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(surface_area))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(surface_area))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+
  labs(x="Raw Material", y="log surface area")
ggsave("fig_retouch_surface_area.png")

##anova
retouched_sa_anova <- aov(log(surface_area) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_retouched_utilized)
summary(retouched_sa_anova)

TukeyHSD(retouched_sa_anova)
```

```{r retouched mass}
#normality check
qqnorm(bpa_retouched_utilized$Mass)
qqline(bpa_retouched_utilized$Mass)
ggplot(bpa_retouched_utilized, mapping = aes(x=Mass))+
  geom_histogram()

#plot 1 
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(Mass))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))

#plot 2
ggplot(bpa_retouched_utilized, aes(x= as.factor(RawMaterial), y = log(Mass))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))
ggsave("fig_retouch_mass.png")

##anova
retouched_mass_anova <- aov(log(Mass) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_retouched_utilized)

TukeyHSD(retouched_mass_anova)
```

### Convergence in Core Shape

```{r}
bpa_cores_updated %>%
  group_by(RawMaterial, SubMember) %>%
  count(CoreBlankSphericity) %>%
  ggplot(aes(x = as.factor(RawMaterial), y = n, fill = factor(CoreBlankSphericity))) +
  coord_flip()+
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Core Shape")+ 
  facet_wrap(~SubMember)
ggsave("fig-core-shape.png")

core_shape_chi_sq <- chisq.test(table(bpa_cores_updated$RawMaterial, bpa_cores_updated$CoreBlankSphericity))
core_shape_chi_sq
```

### Convergence in Core Size

```{r}
#testing form data for normality
qqnorm(bpa_cores_updated$surface_area_cores)
qqline(bpa_cores_updated$surface_area_cores)
ggplot(bpa_cores_updated, mapping = aes(x=surface_area_cores))+
  geom_histogram()

#plot 1
ggplot(bpa_cores_updated, aes(x= as.factor(RawMaterial), y = log(surface_area_cores))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log surface area")
ggsave("fig-core-surface-area.png")

#plot 2
ggplot(bpa_cores_updated, aes(x= as.factor(RawMaterial), y = log(surface_area_cores))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log surface area")
ggsave("fig-core-surface-area-2.png")

#core surface area anova
core_sa_anova <- aov(log(surface_area_cores) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_cores_updated)
summary(core_sa_anova)

TukeyHSD(core_sa_anova)
```

Mass

```{r}
#testing form data for normality
qqnorm(bpa_cores_updated$Mass)
qqline(bpa_cores_updated$Mass)
ggplot(bpa_cores_updated, mapping = aes(x=Mass))+
  geom_histogram()

#plot 1
ggplot(bpa_cores_updated, aes(x= as.factor(RawMaterial), y = log(Mass))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log mass")
ggsave("fig-core-mass.png")

#plot 2
ggplot(bpa_cores_updated, aes(x= as.factor(RawMaterial), y = log(Mass))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log mass")
ggsave("fig-core-mass-2.png")

#core mass anova
##anova
core_mass_anova <- aov(log(Mass) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_cores_updated)

TukeyHSD(core_mass_anova)
```

### Core Reduction Strategy

```{r}
bpa_cores_updated %>%
  group_by(RawMaterial, SubMember) %>%
  count(EvidenceBipolar) %>%
  ggplot(aes(x = as.factor(RawMaterial), y = n, fill = factor(EvidenceBipolar))) +
  coord_flip()+
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Reduction Strategy")+ 
  facet_wrap(~SubMember)
ggsave("fig-core-reduction-strategy.png")

core_reduction_chi_sq <- chisq.test(table(bpa_cores_updated$RawMaterial, bpa_cores_updated$EvidenceBipolarPercussion))
core_reduction_chi_sq
```

### Core Typology

```{r}
bpa_cores_updated %>%
  group_by(RawMaterial, SubMember) %>%
  count(CoreSubTypology) %>%
  ggplot(aes(x = as.factor(RawMaterial), y = n, fill = factor(CoreSubTypology))) +
  coord_flip()+
  geom_col(position = "fill", colour="black")+
  labs(x="Raw Material",y="Proportion", fill="Core Sub Typology")+ 
  facet_wrap(~SubMember)
ggsave("fig-core-typology.png")
```

### **Reduction Intensity**

ARI

```{r}
#average flake length
av_flake_lengths = aggregate(bpa_flakes_updated$TechLength, list(bpa_flakes_updated$RawMaterial, bpa_flakes_updated$SubMember), FUN=mean) 

#average core length
av_core_lengths = aggregate(bpa_cores_updated$MaxLength, list(bpa_cores_updated$RawMaterial, bpa_cores_updated$SubMember), FUN=mean) 

write.csv(av_flake_lengths,file='C:/Users/caleb/Desktop/bpa_data/av_flake_lengths.csv', row.names=FALSE) 
write.csv(av_core_lengths,file='C:/Users/caleb/Desktop/bpa_data/av_core_lengths.csv', row.names=FALSE)
ARI = read.csv("C:/Users/caleb/Desktop/bpa_data/ARI.csv")

#plot
ggplot(data=ARI, mapping = aes(x=as.factor(RawMaterial), y=ARI)) +
  geom_bar(stat="identity") +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1))+
  labs(x="Raw Material", y="ARI")
ggsave("fig-ARI.png")
```

SDI

```{r}
#testing form data for normality
qqnorm(bpa_cores_updated$SDI)
qqline(bpa_cores_updated$SDI)
ggplot(bpa_cores_updated, mapping = aes(x=SDI))+
  geom_histogram()

#plot 1
ggplot(bpa_cores_updated, aes(x= as.factor(RawMaterial), y = log(SDI))) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log SDI")
ggsave("fig-core-SDI.png")

##anova
SDI_anova <- aov(log(SDI) ~ as.factor(RawMaterial) + as.factor(SubMember), data=bpa_cores_updated, group = T)
summary(SDI_anova)
TukeyHSD(SDI_anova)

#plot 2
ggplot(bpa_cores_updated, aes(x= as.factor(RawMaterial), y = log(SDI))) +
  geom_boxplot() +
  facet_wrap(vars(SubMember)) +
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1))+  
  labs(x="Raw Material", y="log SDI")
ggsave("fig-core-SDI.png")
```

## How does this work inform my plans for the next phases of my proposal? (Discussion)
